---
title: "Lab 5"
author: "HaiXiao Lu"
date: "3/8/2021"
output:
  html_document:
    toc: yes
    highlight: zenburn
    toc_depth: 6
    number_sections: yes
    theme: lumen
  pdf_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1 : Make coin-die output using an R funtion

### 1(a). Use cdbbox() to make some useful output

```{r, echo = FALSE, warning = FALSE, message=FALSE}
cdbbox <- function(k = 1, lik, theta, h1="s"){ # k = 1...6
    # xtable for latex output
    if(!require(xtable)) install.packages(xtable)
    # rename the first and second components of the likelihood
    lik1 <- lik[1]
    lik2 <- lik[2]
    # Make a prior that has the desired characteristics
    # if h1 small "s" then ...else...
    ifelse(h1=="s", pi1 <- k/6*lik2/(lik1 + k/6 * lik2), 
           pi1 <- lik2/(lik2 + k/6 * lik1))
    # sum of probalities is 1
    prior <- c(pi1, 1- pi1)
    # lik <- c(lik1, lik2)
    h <- prior * lik
    # Bayes
    post <- h/sum(h)
    # make a matrix for the bayes box
    mat <- cbind(theta, prior, lik, h, post)
    rownames(mat) <- 1:length(lik)
    Totals <- c(NA, sum(prior), NA, sum(h), sum(post))
    mat2 <- rbind(mat, Totals)
    # Make plots 
     layout(matrix(c(1, 2), nr = 1, nc = 2, byrow = TRUE))
    # barplot(matrix(prior, nc=2, nr = 1, byrow = TRUE, 
    #                dimnames = list(c("Coin"), theta)),
    #                ylim =c(0, 1), las = 1, main = "Prior Not the Proposal")
            
    barplot(matrix(c(0.5, 0.5), nc = 2, nr = 1, byrow = TRUE,
                   dimnames = list(c("Coin"), theta)), 
                   ylim = c(0,1),col="olivedrab2",
                   las = 1, main = "Proposal\n Uniform")
    
    barplot(matrix(h, nc = 2, nr = 1, byrow = TRUE,
                   dimnames = list(c("Coin"),theta)), 
                   ylim = c(0, max(h) + 0.5*max(h)), 
                   las = 1,col="red2",
                   main = "Proportional to target\n h")
    
    # barplot(matrix(post, nc=2, nr = 1, byrow = TRUE,
    #                dimnames = list(c("Coin"), theta)),
    #                ylim = c(0,max(post)+0.5*max(post)),
    #                las = 1, col = "green3",
    #                main = "Posterior target /n post")
    # Return a list of useful objects
    return(list(bbox = mat2, latex=xtable(mat2, digits=6),
                mat = mat, h=h, h1=h1, k=k))
    
}
# retrive the function
cdbbox(k=2, lik=dbinom(x=4,size=10, prob=c(0.4, 0.8)),
       theta=c(0.4, 0.8), h1="s") -> result
result$bbox
result$k
```


### 1(a)-iii. Acceptance set

From high to low h values, the acceptance set should be {1, 2, 3, 4} to {1, 2, 3, 4}, 
such as {1, 2}, {1, 3}, {2,3}, etc.

Otherwise, anything that relates to set {5, 6} will be rejected, 
such as {1,5},{1, 6}, {4, 6}, etc.


### 1(b). Improve the graphics from the part(a) function

```{r, echo = FALSE}
cdbbox <- function(k = 1, lik, theta, h1="s"){ # k = 1...6
    # xtable for latex output
    if(!require(xtable)) install.packages(xtable)
    # rename the first and second components of the likelihood
    lik1 <- lik[1]
    lik2 <- lik[2]
    # Make a prior that has the desired characteristics
    # if h1 small "s" then ...else...
    ifelse(h1=="s", pi1 <- k/6*lik2/(lik1 + k/6 * lik2), 
           pi1 <- lik2/(lik2 + k/6 * lik1))
    # sum of probalities is 1
    prior <- c(pi1, 1- pi1)
    # lik <- c(lik1, lik2)
    h <- prior * lik
    # Bayes
    post <- h/sum(h)
    # make a matrix for the bayes box
    mat <- cbind(theta, prior, lik, h, post)
    rownames(mat) <- 1:length(lik)
    Totals <- c(NA, sum(prior), NA, sum(h), sum(post))
    mat2 <- rbind(mat, Totals)
    # Make plots 

                 
    pie(matrix(c(0.5, 0.5), nc = 2, nr = 1, byrow = TRUE,
                   dimnames = list(c("Coin"), theta)), 
                   ylim = c(0,1),col=c("black", "red"), density = 50,
                   las = 1)
                   title(main =list("Proposal\n Uniform", 
                                    col = "red", font = 4))
    
   boxplot(matrix(h, nc = 2, nr = 1, byrow = TRUE,
                   dimnames = list(c("Coin"),theta)), 
                   ylim = c(0, max(h) + 0.5*max(h)), 
                   las = 1)
                  title(main = list("Proportional to target\n h", 
                                    font = 4, col = "red"))
    layout(matrix(c(1, 2), nr = 1, nc = 2, byrow = TRUE))
    
    # Return a list of useful objects
    return(list(bbox = mat2, latex=xtable(mat2, digits=6),
                mat = mat, h=h, h1=h1, k=k))
    
}

# retrive the function
cdbbox(k=2, lik=dbinom(x=4,size=10, prob=c(0.4, 0.8)),
       theta=c(0.4, 0.8), h1="s") -> result
```


## 2. Derive the code snippet using latex

`ifelse(h1=="s", pi1<-k/6*lik2/(lik1+k/6*lik2), pil<- lik2/(lik2+k/6*lik1))`

This line of code means, pi1 will be accepted with probability of (current value) divided (proposed value), else rejected. 

This can be derived as following formula:

`let h1 == "s" ---> min()`
`pi1 ---> aij `
`k/6*lik2 ---> hj`
`lik1+k6*lik2 ---> hi`

$$a_{ij} = min (1, \frac{h_j}{h_i})$$


# Task 2 Make coin-die simulations in R and interpret them

```{r, echo = FALSE}
# This function relies on getthing h vectors that match E2
# Use cdbbox() to get the correct h vector
# Nothice the ... - this corresponds to the ... in the barplot
# you can use additional argument which pass to the barplot
coindie<-function(n=100, h=c(1/4,3/4),E2=c(5,6),init=1,...){
library(xtable)
dieset<-c()
dieset[1]<-"E1"
die<-function(n=1){
sample(1:6,size=n,replace=TRUE)
}

coin<-function(n=1){
sample(1:2,size=n,replace=TRUE)
}
face<-c()
alpha<-c() # holds acceptance probs
alpha[1]<-1
post<-c()# post sample
prop<-c() # vec of proposed states 1s and 2s
prop[1]=init # initial state
post[1]=prop[1]
dice<-c()
dice[1]<-die()

for(i in 2:n){ # starts at 2 because initial value given above
prop[i]<-coin()
alpha[i]=min(1,h[prop[i]]/h[post[i-1]])

dice[i]<-die()
ifelse(alpha[i]==1,dieset[i]<-"E1",dieset[i]<-"E2")
# is x an element of set y
if(alpha[i]==1 | (is.element(dice[i],E2) & alpha[i]!=1)){post[i]<-prop[i]}
else{post[i]<-post[i-1]}
 }  
res<-matrix(c(prop,round(alpha,2),dieset,dice,post ),nc=5,nr=n,byrow=FALSE,dimnames=list(1:n,c("proposal","alpha", "E","dice","post")))
sim<-table(post)/n
postexact<-h/sum(h)
barplot(sim,..., main = "Proportional to target h", col =c( "green3", "red"))
return(list(iter=res,sim=sim,postexact=postexact,post=post,xtable=xtable(res,dig=1)) )
}

# retrive the function
coindie(n = 10, h=c(0.6, 0.4), E2 = c(2, 3, 4, 5))
```

To improve of the graph, I added a title, changed color,width and angle. 

## 2.2 Use of output of cdbbox as input to the coindie()
```{r, echo = FALSE}
coindie<-function(n=100, h=c(1/4,3/4),E2=c(5,6),init=1,...){
library(xtable)
dieset<-c()
dieset[1]<-"E1"
die<-function(n=1){
sample(1:6,size=n,replace=TRUE)
}

coin<-function(n=1){
sample(1:2,size=n,replace=TRUE)
}
face<-c()
alpha<-c() # holds acceptance probs
alpha[1]<-1
post<-c()# post sample
prop<-c() # vec of proposed states 1s and 2s
prop[1]=init # initial state
post[1]=prop[1]
dice<-c()
dice[1]<-die()

for(i in 2:n){ # starts at 2 because initial value given above
prop[i]<-coin()
alpha[i]=min(1,h[prop[i]]/h[post[i-1]])

dice[i]<-die()
ifelse(alpha[i]==1,dieset[i]<-"E1",dieset[i]<-"E2")
# is x an element of set y
if(alpha[i]==1 | (is.element(dice[i],E2) & alpha[i]!=1)){post[i]<-prop[i]}
else{post[i]<-post[i-1]}
 }  
res<-matrix(c(prop,round(alpha,2),dieset,dice,post ),nc=5,nr=n,byrow=FALSE,dimnames=list(1:n,c("proposal","alpha", "E","dice","post")))
sim<-table(post)/n
postexact<-h/sum(h)
barplot(sim,..., main = "Proportional to target h", col =c( "lightblue", "mistyrose"),
        angle = 90, width = c(5, 10))
return(list(iter=res,sim=sim,postexact=postexact,post=post,xtable=xtable(res,dig=1)) )

}

result2 <- coindie(n = 200, h = result$h, E2 = 1:result$k)

```

I used h values which is `prior * likelihood` from cdbbox() function. 
Also set E2 to 1:k value (2 stage) from cdbbox() function output. 


# Task 3. Make a simulation with any number of discrete theta values

## 3.1 Explain the code snippet

Explanation: first select a random value (U) from uniform(1). 
If the value in uniform is less than the value in alpha(alpha[i]), 
then we accept the proposal ({post[i] < prop[i]}).
Otherwise, we reject the proposal. Posterior value is just the same one as previous. 


## 3.2 Using a uniform prior to create a simulated posterior histogram
```{r, echo = FALSE}
simR <- function(n, init = 0.5, h=c(0.03344302, 0.06165627)){
alpha<-c() # holds transition probs
alpha[1]<-1
u<-c() # holds uniform values
u[1]<-1
post<-c()# post sample
prop<-c() # vec of proposed states 1s and 2s
prop[1]=1 # initial state
post[1]=prop[1]
for(i in 2:n){ # starts at 2 because initial value given above
# proposal state 
prop[i]=sample(1:length(h),1,replace=TRUE)
# calculate alpha
# notice h[prop[i]] gives the actual value of h
alpha[i]=min(1,h[prop[i]]/h[post[i-1]])
# to calculate accepting proposal with prob alpha
# select a random value from  a uniform (0,1)
u[i]=runif(1)
if(u[i]<=alpha[i]){post[i]<-prop[i]}
else{post[i]<-post[i-1]}
}
res<-matrix(c(prop,u,alpha,post ),nc=4,nr=n,byrow=FALSE)
sim<-table(post)/n

barplot(sim)
plot(sim)
postexact<-h/sum(h)
# The returned output is a list 
# Use obj$ to obtain whatever interests you
return(list(iter=res,sim=sim,postexact=postexact,post=post)) -> a
}
simR(40)
a
hist(a$post, freq = FALSE)
```


## 3.3 Improve the graphical output by editing the function


# Task 4. Use different proposals
```{r, echo = FALSE}
# Again for the discrete case
simRQ<-function(n=1000,init=1, h=c(1,1),pr=c(1,1)/2,...){
alpha<-c() # holds transition probs
alpha[1]<-1
u<-c() # holds uniform values
u[1]<-1
post<-c()# post sample
prop<-c() # vec of proposed states 1s and 2s etc
prop[1]=init # initial state
post[1]=prop[1]
q<-function(x){pr[x]}
for(i in 2:n){ # starts at 2 because initial value given above
#make a sample from the proposal
sample(1:length(h),1,replace=TRUE,prob=pr)->prop[i]
#Calculate alpha adjusting for the proposal being non uniform
alpha[i]=min(1,h[prop[i]]*q(post[i-1])/(h[post[i-1]]*q(prop[i])))
# now choose the proposal with probability alpha
u[i]=runif(1)
if(u[i]<=alpha[i]){post[i]<-prop[i]}
else{post[i]=post[i-1]}
}
res<-matrix(c(prop,u,alpha,post ),nc=4,nr=n,byrow=FALSE,dimnames=list(1:n,c("prop","u","alpha","post")))
sim<-table(post)/n
postexact<-h/sum(h)
barplot(sim,...)
hist(post, freq = FALSE)
tmp<-c()
ifelse(length(res[,1])>=20,tmp<-res[1:20,],tmp<-res)
return(list(iter=tmp,sim=sim,post=postexact) )
}
# not sure what pr is 
simRQ(1000, h = c(3, 1), pr = c(1, 2)/3) -> ans1
hist(ans1$iter[,3], freq = FALSE)

```


# Task 5. Make simulations from a continuous paramter with any proposal
```{r}
simRC<-function(n=10,init=0.5, h=function(theta){dunif(theta)*dbinom(x=4,size=10,prob=theta)},a=3,b=4){
#dbeta(x, shape1, shape2, ncp = 0, log = FALSE)
alpha<-c() # holds transition probs
alpha[1]<-1
u<-c() # holds uniform values
u[1]<-1
post<-c()# post sample
prop<-c() # vec of proposed states 1s and 2s
prop[1]=init # initial state
post[1]=prop[1]
q=function(x){dbeta(x,a,b)}
for(i in 2:n){ # starts at 2 because initial value given above
rbeta(1,a,b)->prop[i]

alpha[i]=min(1,h(prop[i])*q(post[i-1])/(h(post[i-1])*q(prop[i])))
u[i]=runif(1)
ifelse(u[i]<=alpha[i],post[i]<-prop[i],post[i]<-post[i-1])
}
res<-matrix(c(prop,u,alpha,post ),nc=4,nr=n,byrow=FALSE,dimnames=list(1:n,c("prop","u","alpha","post")))
hist(post, freq=FALSE)
return(list(matrix=res,summary=summary(post)) )
}

simRC(10000)
```


# Task 6. Use JAGS to carry out a Gibbs sample for SLR

